\documentclass{res}

% suppress widows/orphans:
\widowpenalty=1000
\clubpenalty=1000
\usepackage{helvetica} % uses helvetica postscript font
\usepackage{hyperref}
\hypersetup{
    pdfstartview={FitH},   % fits the width of the page to the window
    pdftitle={Resume},
    pdfauthor={Yaroslav Lyakh},
    pdfkeywords={CV} {Java} {multithreading} {RDBMS} {multithreading}, %list of keywords
    colorlinks=true,      % false: boxed links; true: colored links
    % linkcolor=black,          color of internal links (change box color with linkbordercolor)
    % citecolor=green,        color of links to bibliography
    % filecolor=magenta,      color of file links
    urlcolor=black          % color of external links
}
\usepackage{needspace}
\newsectionwidth{0pt}  % So the text is not indented under section headings

\usepackage{scalefnt}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\itemsep=0in
\parsep=0in
\usepackage{enumitem}
\setlist[itemize]{itemsep=-2pt, topsep=4pt, labelindent=\parindent}

% hints to the word-breaking algo
\hyphenation{conti-nuous}

\newcommand{\emp}[1]{{\bf#1}}

\newcommand{\jobdes}[4]{\needspace{3\baselineskip} %to keep with next paragraph
{\noindent \bf #3\hspace{2ex}}{{\em \small #1}}{\hfill #2}\\{{\it \small #4.}}}

\newcommand{\aggjobdes}[4]{\needspace{3\baselineskip} %to keep with next paragraph
{\noindent \bf #1\hspace{2ex}}{\hfill #2}\\
{{\noindent \small \textit{ \textbf{ Companies:} {\hfill #3}}}}\\{{\it \small #4.}}}

\newcommand{\osection}[1]{\section{\sc {\Large \textbf{#1}\\}} \vspace{0.30cm}}

% set formatting, put semicolon and dot as needed
\newcommand{\profileItem}[2]{\item {\sl #1:} #2.}
% \newcommand{\profileItem}[2]{\item {\sc #1:} #2.}

% \sl is slanted, not \it (italic) !!! see also http://www.combinatorics.net/weblib/options/style.html

\usepackage{fancyhdr} % use this package to get a 2 line header
\pagestyle{fancy}     % set pagestyle for document
\renewcommand{\headrulewidth}{0pt} %suppress line drawn by default by fancyhdr
\setlength{\textheight}{10in} %increase text height to fit on 1-page
% \textwidth=5.2in % increase textwidth to get smaller right margin
\setlength{\headsep}{4pt}  % space between header and text
\setlength{\headheight}{12pt} % allow room for header
\fancyheadoffset{1cm}
\rhead{ {\it Yaroslav Lyakh (\href{mailto:l_yara@bigmir.net}{L\_yara@bigmir.net}; tel: +447501 301536)}} %put text in header (right side)
% \rhead{ {\it  \thepage}} put text in header (right side)
\cfoot{}                              %       the foot is empty
\topmargin=-2cm %start text higher on the page
\footskip=4cm

\setlength{\parindent}{2em} % paragraph indent

% switch full / short version
% see http://tex.stackexchange.com/questions/1492/passing-parameters-to-a-document for detail
% use  pdflatex "\def\longversion{1} \input{Y_Lyakh.tex}"
% command use pandoc for exporting:
% pandoc Yaroslav_Lyakh.tex -o Yaroslav_Lyakh.docx
\begin{document}
% \thispagestyle{empty} %this page has no header
% \name{Yaroslav Lyakh}
\begin{resume}

%\osection{Summary}
% \vspace{0.20cm}
\indent Team Lead / Senior Software Engineer / Tech Lead with over eighteen years of experience
% in a variety of roles
across a number of industries: Finance, Banking, Web Security, Oil~\&~Gas, Online Gambling, Utilities and Public Safety. Strong engineering background, working knowledge and proven skills in all stages of the software development life~cycle. Almost 6~years of team and project leading experience, ability to interact well at all levels and articulate the IT-specific issues to non-IT colleagues in a clear business-transparent way.
% \indent Looking for a challenging role where my skills and qualifications will add value and make a significant contribution to the company.
% \indent More than ten years of Java development experience; extensive involvement in high-load distributed low-latency data processing systems development for last five years.

\osection{Career Highlights}
\begin{itemize}
\item \emp{Formed and led the Software Development team} (up to eight people) in a multi-million GIS MT project for Gazprom gas-transportation giant. Our (VTG team) role in this project constantly increased due to the high level of technical and business expertise developed by team members combined with collaborative rapport built with our partners.
\item Introduced a number of system design improvements, particularly enabling a \emp{massive multi-threading data processing} on top of Spring Integration infrastructure with Fizzback team. The scalability level is now sufficient for any foreseeable load.
\item Designed and developed the \emp{data collection, verification and migration infrastructure} for Utility and Oil \& Gas - specific databases (both spatial and attribute parts). The changes in the business process proposed by our team and backed by our software enabled a productivity increase in 3-4 times compared to old approach.
\item Holding two \emp{patents} awarded on IT innovations in \emp{Data Presentation} and \emp{Data Visualisation} areas.
\item As a \emp{Principal Developer / Architect} made a key contribution to the success of the PS\&E project when worked for Kring LLC. Despite unusual team configuration and inconvenient technology set the system showed \emp{good performance and reliability} achieved with a \emp{purpose-built testing facility}.
\item Built a set of the state- and local- budget \emp{data management tools} when worked in a large KPMG consultancy projects for the Ministry of Finance of Ukraine. The projects enjoyed the great success and were credited with eliminating 2 months from the budget development cycle.
\item Two times won the special prizes in \emp{``Thinking Out of the Box''} nominations during the traditional \emp{Hack Day} events in MapInfo / Pitney Bowes.
\end{itemize}

% disable for now
\iffalse
\osection{Technical Profile}
% \vspace{0.20cm}
\begin{itemize}
  \profileItem {Full software development lifecycle (SLDC)} {requirement gathering and analysis, design, prototyping, implementation, performance tuning}
  \profileItem {Development process organisation} {\emp{Scrum} (and other \emp{Agile} flavours), RUP, Waterfall. Team work planning and coordination, peer code reviews, team members coaching and mentoring. \emp{Continuous Integration (CI)} with Jenkins}
  \profileItem {JVM - based languages} {\emp{Java} (from 1.3 to 8.0) SE/EE with \emp{multi-threading, collections} and \emp{synchronisation primitives} (lock/barrier/phaser, Guava Monitor). Basic knowledge of \emp{Scala} and \emp{Groovy}; currently studying \emp{Clojure}}
  \profileItem{Performance diagnosis and optimisation} {performance analysis with \emp{jMeter, VisualVM} and \emp{JMH}. Multi-threading systems development; runtime performance tuning incl. Java GC optimisation}
  \profileItem {Relational Databases (RDBMS)} {\emp{Oracle, MySQL, MS SQL Server;} complex \emp{SQL} query authoring and optimisation; \emp{PL/SQL} stored procedures development and performance tuning}
  \profileItem{Enterprise Technologies} {Enterprise Integration Patterns (EIP), Service Oriented Architecture (SOA), \emp{messaging, caching}. Server-side and middleware development; REST-based microservices implementations; Raft distributed consensus algorithm}
  \profileItem{Data Modelling, Data Analysis and Presentation, Extract-Transform-Load (ETL)} {data denormalisation, complex \emp{data transformation and cleansing}; advanced \emp{report building}}
  \profileItem{Other Languages} {\emp{C\#}, some \emp{Python}, \emp{R/Octave}. Shell Script, XML/XSL, JSON, \LaTeX, MD markup}
  \profileItem{Operation Systems, Tools, Technologies, Libraries} {\emp{Linux, Mac OS} and \emp{Windows}. Apache (Ant, Commons Lang, Commons IO, Log4j, \emp{Maven}, \emp{Tomcat}). Crystal Reports. Google (Guava, \emp{Mockito}). JDBC, Hibernate ORM. jBPM/Activiti. Java Swing. \emp{jUnit}, EasyMock, jMock. \emp{Spring} (Core, Integration, MVC, Roo). IDEs: IntelliJ IDEA, Eclipse, Visual Studio. Source Code Control with \emp{Git} and \emp{SVN/Subversion}. \emp{UML} with Sparx Enterprise Architect and Rational Rose. Emacs}
\end{itemize}
\fi

\osection{Experience}
\aggjobdes {Senior Software Engineer / Senior Java Developer}{December 2007 - Present}
{Betfair, ION Trading, Fizzback (NICE Co), ScanSafe (Cisco), MapInfo: UK; Original~Solutions: Ireland.}
{Low-latency high-throughput Java 6, 7 and 8 - based systems. Oracle SQL and PL/SQL. Zookeeper, in-house Raft consensus algorithm implementation. REST-based microservices. Continuous Integration (CI) with Maven, Jenkins and Go CD. TDD (jUnit / TestNG + Mockito), Acceptance Testing (Robot Framework and purpose-built frameworks). Performance analysis with jMeter, VisualVM and JMH; runtime performance tuning including Java GC optimisation. Groovy, Spring (Core, Spring Integration, SpEL, Spring Roo), Hibernate, Activiti BPM (jBPM). GIS, ETL. Git, Subversion. .NET (C\#). Agile (Scrum), peer reviews, mentoring junior team members.}
\begin{itemize}
 \item Developed new functionality for the large-scale high-throughput modules like One-Way Matching for Betfair, Communication Infrastructure for Fizzback or Data Analysis and Reporting for ScanSafe.
 \item Heavily involved in so-called ``parallelising'' works: redesign and re-implementation of the performance - critical parts as the distributed or at least multithreaded modules: Multiple Orchestrator project in Betfair, Parallel Query Executor in ION Trading and on-the-fly data processing module in Fizzback.
 \item In Betfair, took a major part in transition of the legacy Oracle-based ``traditional'' (i.e. built on classical read-process-write cycle) Matcher implementation into more robust stream-based architecture based on Raft distributed consensus algorithm. Lead a similar project in Fizzback where we transformed an existing project into much more scalable and reliable messaging - based solution.
 \item As part of the performance improvement works designed and implemented the project-specific Performance Testing facilities for Betfair, Fizzback, ScanSafe and MapInfo.
 \item Maintained a consistent attention on the operational characteristics of the projects. In ION Trading, implemented the Attribute Replacement feature which greatly increased the system’s configuration flexibility and made administering and configuring of the whole product much more convenient. In Fizzback, deeply re-developed the Communication and Reporting sub-modules to make configuration and maintenance more convenient; these changes (like specialised configuration DSL and flexible report templates) greatly improved overall system's maintainability and customer provisioning time.
 \item Enhanced the Continuous Integration infrastructure of all projects I've been involved in. In ION Trading, re-implemented a project-specific Maven plugins and streamlined the existed Robot Framework modules increasing the test reliability and reducing the test execution time. In Fizzback, improved the execution time of the whole Unit Testing process (mostly by database-related optimisation).
 \item Provided training and on-the-fly consultations to the junior team members. In Fizzback, developed good working relations with colleagues from Operations, QA and Customer Support departments; combined with good ``pure'' development work this decreased an average defect response time and improved overall IT team flexibility and confidence.
 \end{itemize}

\aggjobdes {Team Lead / Principal Software Engineer / IT Department Head}{January 2002 - December 2007}
{Kring LLC: Ukraine/Denmark; VTG (Vnipitransgaz), Privatinvest Retail Bank: Ukraine}
{Java 1.3 and 5, multi-threading, GIS, Oracle Spatial Data Option (SDO). Bentley MicroStation and GeoGraphics, Intergraph suite, Spring, O/R Mapping, Swing, JDBC, jUnit, jMock, Eclipse, Subversion, Oracle 8i, 9i, and 10g: complex SQL query authoring and optimisation, PL/SQL stored procedures development and performance tuning. Data Modelling, Data Analysis and Presentation, Extract-Transform-Load (ETL); data denormalisation, complex data transformation and cleansing; advanced report building.  UML, Sparx EA, Rational Rose, XML / XSLT. Scrum and RUP. TETRA – based mobile devices communication}
\begin{itemize} % \itemsep -2pt % reduce space between items
 \item In Kring LLC, lead the offshore development project of a major Public Safety and Emergency (PS\&E) system development. The project consisted of the server part built on Java (conversion of existing military Command and Control Information System) and mobile .Net Compact Framework - based clients. Modules for interacting with on-board GPS receivers and specialised communication hardware were implemented along with purpose-built integration testing facility.
 \item In VTG, formed and led the Software Development Team involved in the major distributed Geographical Information System (GIS MT) for the Gasprom gas transportation company. My role included business process analysis and modelling, requirement analysis, system design and architecting in cooperation with customer’s experts and other development teams and work coordination with partners.
 \item Inside VTG team was responsible for planning development efforts for team members, technical oversight and coaching as well as hands-on project work. Two modules developed by our team were honoured with special patents of the Russian Federation. The project has exceeded the customer’s expectation and was recommended for adoption in all Gasprom child companies.
 \item In Privatinvest lead the technical implementation of the major expansion of the bank from single office to 22 offices distributed around the country. Created and expanded the IT department, introduced a number of innovations that allowed us to complete the technical tasks in only 5 months – a month ahead of original schedule.
\end{itemize}

\aggjobdes {Software Developer / Analyst Programmer}{November 1996 - December 2001}
{Lost Boys: The Netherlands; KPMG: Consulting Project in the Ministry of Finance of Ukraine}
{Java 1.3, multi-threading, O/R Mapping, Swing, JDBC, jUnit, Eclipse, Subversion. MS SQL Server, MS Office (particularly MS Access), VBA, Delphi, Crystal Reports, IIS + ASP, XML/XSLT}
\begin{itemize} % \itemsep -2pt % reduce space between items
 \item In Lost Boys, was selected to be part of the special fine-tuning and performance optimization team – a distinction given to the best developers at the Company.
 \item In KPMG, took a first-hand part in a planning, design and implementation of various projects devoted to the budgeting process of the state and local budgets.  The projects were credited with eliminating 2 months from the budget development cycle and greatly expanded the analytic capabilities of the Ministry of Finance of Ukraine.
\end{itemize}

\osection{SPECIALS}
\begin{itemize}
 \item Profiles: StackOverflow \url{http://stackoverflow.com/users/900174/l-yara}, GitHub \url{http://github.com/l-yara}, LinkedIn \url{http://uk.linkedin.com/in/ylyakh}.
 \item Basic knowledge of Scala, Python, R, Shell Script, XML/XSL, JSON and \LaTeX; currently studying Clojure. An active member of Data Science London, London Java Community (LJC) and London Clojurians communities.
 \item
   %% Strong interest in Functional Programming and Artificial Intellect (AI) / Business Intelligence (BI) system development.
   Took a number of online courses on Finance (including Financial Markets and Computational Investing), Computer Science (Algorithms, Functional and Reactive Programming) and Data Science (incl. Big Data, Machine Learning and Artificial Intelligence) - see full list in LinkedIn profile.
 \item Education: Engineer Diploma (USSR equivalent of MSc) with Honours in Electrical Engineering from National Technical University of Ukraine (KPI) in Kiev, Ukraine, 1996.
 \item Sports: volleyball (former player of ``CBPhoto'' Irish Division 1 team, now playing in Kent amateur league), sailing (former pupil of Kiev Olympic School).
\end{itemize}


\end{resume}
\end{document}
