\documentclass{res}

% suppress widows/orphans:
\widowpenalty=1000
\clubpenalty=1000
\usepackage{helvetica} % uses helvetica postscript font
\usepackage{hyperref}
\hypersetup{
    pdfstartview={FitH},   % fits the width of the page to the window
    pdftitle={Resume},
    pdfauthor={Yaroslav Lyakh},
    pdfkeywords={CV} {Java} {multithreading} {RDBMS} {Team Lead}, %list of keywords
    colorlinks=true,      % false: boxed links; true: colored links
    % linkcolor=black,          color of internal links (change box color with linkbordercolor)
    % citecolor=green,        color of links to bibliography
    % filecolor=magenta,      color of file links
    urlcolor=black          % color of external links
}
\usepackage{needspace}
\newsectionwidth{0pt}  % So the text is not indented under section headings

\usepackage{scalefnt}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\itemsep=0in
\parsep=0in
\usepackage{enumitem}
\setlist[itemize]{itemsep=-2pt, topsep=4pt, labelindent=\parindent}

\newcommand{\jobdes}[4]{\needspace{3\baselineskip} %to keep with next paragraph
{\noindent \bf #3\hspace{2ex}}{{\em \small #1}}{\hfill #2}\\{{\it \small #4.}}}

\newcommand{\osection}[1]{\section{\sc {\Large \textbf{#1}\\}} \vspace{0.30cm}}

% set formatting, put semicolon and dot as needed
\newcommand{\profileItem}[2]{\item {\sc #1:} #2.}
\newcommand{\emp}[1]{{\bf#1}}

\usepackage{fancyhdr} % use this package to get a 2 line header
\pagestyle{fancy}     % set pagestyle for document
\renewcommand{\headrulewidth}{0pt} %suppress line drawn by default by fancyhdr
\setlength{\textheight}{10in} %increase text height to fit on 1-page
% \textwidth=5.2in % increase textwidth to get smaller right margin
\setlength{\headsep}{4pt}  % space between header and text
\setlength{\headheight}{12pt} % allow room for header
\fancyheadoffset{1cm}
\rhead{ {\it Yaroslav Lyakh (\href{mailto:l_yara@bigmir.net}{L\_yara@bigmir.net}; tel: +447501 301536)}} %put text in header (right side)
% \rhead{ {\it  \thepage}} put text in header (right side)
\cfoot{}                              %       the foot is empty
\topmargin=-2cm %start text higher on the page
\footskip=4cm

\setlength{\parindent}{2em} % paragraph indent

% switch full / short version
% see http://tex.stackexchange.com/questions/1492/passing-parameters-to-a-document for detail
% use  pdflatex "\def\longversion{1} \input{Yaroslav__Lyakh.tex}"
% command use pandoc for exporting:
% pandoc Yaroslav_Lyakh.tex -o Yaroslav_Lyakh.docx
\newif\ifFullVersion
%\FullVersiontrue % un-comment for the ``full'' version
% \ifdefined\longversion
%   \FullVersiontrue  un-comment for the ``full'' version
% \else
%   \FullVersionfalse
% \fi

\begin{document}
% \thispagestyle{empty} %this page has no header
% \name{Yaroslav Lyakh}
\begin{resume}

%\osection{Summary}
% \vspace{0.20cm}
\indent Senior Software Engineer / Tech Lead with over eighteen years of experience across a number of industries: Finance, Banking, Web Security, Oil~\&~Gas, Online Gambling, Utilities and Public Safety. \\
\indent More than five~years of Team Leading experience, proven ability to lead by example, coach / mentor junior team members and cooperate with non-IT colleagues.\\
\ifFullVersion
Strong engineering background, working knowledge and proven skills in all stages of the software development life-cycle. \\
\indent Looking for a challenging role where my skills and qualifications will add value and make a significant contribution to the company.
\else
\indent Recent Java development experience for 10+ years; distributed low-latency massive data processing systems development for last five years.
\fi

\ifFullVersion
\osection{Career Highlights}
\begin{itemize}
\item \emp{Formed and led the Software Development team} (up to eight people) in a multi-million GIS MT project for Gazprom gas-transportation giant. Our role in this project constantly increased due to the high level of technical and business expertise developed by team members combined with collaborative rapport built with our partners.
\item Introduced a number of system design improvements, particularly enabling a \emp{massive multi-threading data processing} on top of Spring Integration infrastructure with Fizzback team. The scalability level is now sufficient for any foreseeable load.
\item Designed and developed the \emp{data collection, verification and migration infrastructure} for Utility and Oil \& Gas - specific databases (both spatial and attribute parts). The changes in the business process proposed by our team and backed by our software enabled a productivity increase in 3-4 times compared to old approach.
\item Holding two \emp{patents} awarded on IT innovations in \emp{Data Presentation} and \emp{Data Visualisation} areas.
\item As a \emp{Principal Developer / Architect} made a key contribution to the success of the PS\&E project when worked for Kring LLC. Despite unusual team configuration and inconvenient technology set the system showed \emp{good performance and reliability} achieved with a \emp{purpose-built testing facility}.
\item Built a set of the state- and local- budget \emp{data management tools} when worked in a large KPMG consultancy projects for the Ministry of Finance of Ukraine. The projects enjoyed the great success and were credited with eliminating 2 months from the budget development cycle.
\item Two times won the special prizes in \emp{``Thinking Out of the Box''} nominations during the traditional \emp{Hack Day} events in MapInfo / Pitney Bowes.
\end{itemize}
\fi

\osection{Technical Profile}
\vspace{0.20cm}
\begin{itemize}
  \profileItem {Full software development lifecycle (SLDC)} {requirement gathering and analysis, design, prototyping, implementation, quality assurance (QA) organisation, performance testing and tuning}
  \profileItem{Development process organisation} {\emp{Agile} (\emp{Scrum} and XP), RUP, Waterfall. Team work planning and coordination, peer code reviews, team members coaching and mentoring. \emp{Test-Driven Development (TDD), Continuous Integration (CI)} and \emp{Release Management}}
  \profileItem {JVM - based languages} {\emp{Java} (from 1.3 to 8) SE/EE. Skills in \emp{Clojure}, \emp{Scala} and \emp{Groovy}}
  \profileItem{Performance diagnosis and optimisation} {high-load \emp{parallel}, \emp{distributed} and \emp{cloud}-based \emp{low-latency} development. Performance analysis and tuning with \emp{VisualVM, jMeter}}
  \profileItem {Relational Databases (RDBMS)} {\emp{Oracle, MySQL, MS SQL Server;} complex \emp{SQL} query authoring and optimisation; \emp{PL/SQL} stored procedures development and performance tuning}
  \profileItem{Enterprise Technologies} {Enterprise Integration patterns (EIP), Service Oriented Architecture (SOA), \emp{messaging, caching}. Server-side and middleware development; REST-based microservices implementations; Raft distributed consensus algorithm}
  \profileItem{Data Modelling, Data Analysis and Presentation, Extract-Transform-Load (ETL)} {including data denormalisation, complex \emp{data transformation and cleansing}; advanced report building. Holding two patents awarded on IT innovations in Data Presentation and Data Visualisation areas}
\ifFullVersion
Strong interest in \emp{functional programming}, Data Science and Machine Learning / Business Intelligence systems.
\fi
  \profileItem{Other Languages} { \emp{JavaScript} (NodeJS, AngularJS), \emp{C\#}, \emp{Python}, \emp{R/Octave}. Shell Script, XML/XSL, HTML5 / CSS3, JSON, \LaTeX ~typesetting, MD markup, Velocity}
  \profileItem{Tools, Technologies, Libraries} {\emp{Linux, Mac OS} and \emp{Windows}. Apache (Ant, Commons Lang, Commons IO, \emp{Maven}, \emp{Tomcat}); Atlassian (\emp{Confluence}, \emp{Jira}); NoSQL: \emp{MongoDB}; Crystal Reports; Google (Guava, \emp{Mockito}); \emp{Hibernate}; JDBC; jBPM/Activiti; Java \emp{multi-threading, collections} and \emp{synchronisation primitives}. Java Server Pages (JSP); Java Swing; \emp{Jenkins CI}; jMock; \emp{jUnit}; \emp{Spring} (Core, Integration, MVC, Roo). IDEs: IntelliJ IDEA, Eclipse, Visual Studio. \emp{VisualVM} diagnostic tool. Transport (TCP, UDP) and application (HTTP) level network protocols. \emp{Git} and \emp{SVN/Subversion}. \emp{UML}: Sparx Enterprise Architect and Rational Rose. Emacs}

\end{itemize}

\osection{Career History}
\jobdes{Betfair: London, UK} {March 2015 - March 2016}{Senior Java Consultant}
{Low-latency high-throughput systems based on low-level Java 7 and 8 multi-threading primitives (locks/barriers/phaser, Guava Monitor) and concurrent collections. Oracle SQL and PL/SQL. Zookeeper. Raft distributed consensus algorithm. Continuous Integration (CI) with Jenkins and Thoughtworks Go. TDD (TestNG + Mockito), Maven, Spring, Git, Scrum, peer reviews, mentoring junior team members}
\begin{itemize}
 \item In an Agile environment, completed transition of the legacy Oracle PL/SQL and Java 7 - based ``traditional'' (i.e. based on classical read-process-write cycle) Bet Matcher implementation into a Java 8 stream - based architecture built on Raft.
 \item Heavily involved in so-called ``Multiple Orchestrator'' sub-project enabling a significant increase in the degree of parallelism for the whole Bet Matching system.
 \item Designed and implemented a project-specific Performance Testing facility.
 \item Provided coaching and on-the-fly consultations to the junior team members.
 \end{itemize}

\jobdes{ION Trading: London, UK} {July 2013 - March 2015}{Senior Software Developer}
{Java 7 and 8, distributed multithreaded low-latency systems. Continuous Integration (CI) with Jenkins, Behaviour (Acceptance) Testing with the Robot Framework. Java Remote Method Invocation (RMI), TDD (jUnit + Mockito), VisualVM, Maven, Git, Groovy, HTML5, CSS3, Velocity, Scrum, Subversion, DevOps, peer reviews}
\begin{itemize}
 \item Re-designed the Query Execution module as a distributed parallel executing subsystem.
 \item Enhanced the Continuous Integration infrastructure, re-implemented a project-specific Maven plugins. Streamlined the existed Robot Framework modules which increased the test reliability and reduced the test execution time. Improved the Robot test spec HTML/CSS format for better structure and readability.
 \item Implemented the Attribute Replacement feature which greatly improved the system’s configuration flexibility and made administering and configuring of the whole product much more convenient.
 \end{itemize}

\jobdes{Fizzback (now NICE Co): London, UK} {June 2011 - July 2013}{Senior Fullstack Engineer / Java DevOps}
{Java 6 and 7, multi-threading, Spring (Core, Spring Integration, SpEL, Spring Roo), Hibernate, Activiti BPM, Groovy, jUnit, Mockito, VisualVM, Maven, Domain-Specific Languages (DSLs), Python (Jython), HTML 5, CSS 3, JavaScript (NodeJS, AngularJS), SMS-based mobile communicating, Bash,  Mac OS, Linux, Scrum, peer reviews}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Refactored and enhanced a set of modules enabling a massive multi-threading data processing on top of Spring Integration pipelines.
\ifFullVersion
While for some modules it was enough to re-code the critical pieces with the standard “stateless or immutable” approach, the main part of existed system required deep re-design based on Enterprise Integration Patterns (EIP) messaging-based architecture and extensive use of Java multi-threading tools.
\fi
The changes in performance and resource consumption were constantly monitored using jMeter and VisualVM.
 \item Redeveloped the Reporting and Database Access layers, got rid of locking-related problems in both.
\ifFullVersion
Some modules were completely re-built until the required level of performance and reliability was achieved.
\fi
 \item In addition to the scalability improvements deeply re-developed the Communication and Reporting sub-modules to make them more convenient in configuration and maintenance.
\ifFullVersion
   These changes (like introducing ad-hoc configuration DSL and flexible report templates) greatly improved overall system's maintainability and customer provisioning time.
\fi
Improved the execution time of the whole Unit Testing process (mostly by database-related optimisation).
\ifFullVersion
Initiated use of CI service (Jenkins) as a key component of the team’s Scrum Agile methodology approach.
\fi
\item Actively participated in front-end HTML/CSS/JavaScript development and maintenance.
\item Developed good working relations with colleagues from Operations, QA, and Customer Support departments. Combined with good ``pure'' development work this decreased an average defect response time and improved overall IT team flexibility and confidence.
\end{itemize}

\jobdes{ScanSafe Inc (now part of Cisco): London, UK}{July 2010 - June 2011}{Senior Software Developer}
{Java 6, multi - threading, Spring, Hibernate, Tomcat, JSP, Spring MVC, Groovy, jUnit, Mockito, XML/XSLT, jMeter, Hudson, Oracle 11g, Oracle SQL, PL/SQL, VerticaDB, Linux, Sparx EA, Selenium, Agile Development (Scrum), DevOps: Bash (awk, grep), VMWare, TCP and HTTP traffic analysis and optimisation}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Designed and implemented various parts of a major Web Security system based on a set of parallel running services being executed in a private cloud environment.
 \ifFullVersion
   This included a lot of re-design work, particularly switching from the standard Java objects to the immutable structures in combination with elements of functional-oriented programming.
\fi
Data structures like tries and algorithm tricks like specifically built hash functions were used to provide the required scalability level.
 \item Implemented Data Analysis and Reporting modules with sophisticated “on-the-fly” SQL query building. Provided the serious database performance tuning including SQL query re-writing, PL/SQL programming and index optimisation.
\end{itemize}

\jobdes{Pitney Bowes (MapInfo): Chatham, UK}{July 2008 - April 2010}{Senior Software Engineer}
{.NET (C\#), NUnit, Rhino Mocks, Visual Studio, MS Entity Framework, LINQ, WPF/XAML, Web Services, Subversion, Oracle 11g, SQL, PL/SQL, GIS, ETL, Agile Scrum Development}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Took a key part in major system redesign and redevelopment works managed around the Scrum methodology. Innovative approaches like LINQ and SaaS were also featured in this project together with some elements of functional-based design which allowed to cut down the codebase size; the stability and reliability of the whole system also improved. Suggested and implemented a number of innovations like Finite-State Machine formalisation of business logic and BPM workflow.
\end{itemize}

\jobdes {Original Solutions: Limerick, Ireland}{December 2007 - May 2008}{Consulting Software Developer}
{Java 5, multi – threading, Spring, Hibernate, jBoss jBPM, jUnit, Eclipse, Perforce, Ant, Oracle, SQL, PL/SQL}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Provided performance optimization works; created a project-specific performance testing subsystem. Amending jBPM for use in multi-threaded environment was a massive effort but we managed to provide a high level of scalability on our system.
\end{itemize}

\jobdes {Kring LLC: Ukraine - Denmark}{December 2006 - December 2007}{Principal Software Developer/Architect}
{Java 5, Spring, JDBC, jUnit, jMock, Eclipse, Subversion, Ant, Oracle, SQL, .NET Compact Edition (C\#), NUnit, Rhino Mocks, Visual Studio, GIS, TETRA – based mobile devices communication, TCP/UDP protocols}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Lead the offshore development project of a major Public Safety and Emergency (PS\&E) system development. The project consisted of the server part built on Java (conversion of existing military Terma FlexCCIS Command and Control Information System) and mobile clients based in .Net Compact Framework (C\#). Modules for interacting with on-board GPS receivers and specialised communication hardware were implemented along with purpose-built integration testing facility.
\end{itemize}

\jobdes{Vnipitransgaz: Kiev, Ukraine}{July 2002 - November 2006}{Team Lead}
{Java 1.3 and 5, multi-threading, GIS, Oracle Spatial Data Option (SDO), Bentley MicroStation, Intergraph suite, Spring, O/R Mapping, Swing, JDBC, jUnit, jMock, Eclipse, Subversion, Oracle 8i, 9i, and 10g, SQL, PL/SQL, UML, Sparx EA, Rational Rose, XML, MS Office + VBA, XSLT, TDD and RUP}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Formed and led the Software Development Team involved in the major distributed Geographical Information System (GIS MT) for the Gasprom gas transportation company. My role included business process analysis and modelling, requirement analysis, system design and architecting in cooperation with customer’s IT department and other development teams plus work coordination with partners.
 \item Inside team (up to eight people) planned development efforts for team members and worked on the project hands-on; provided technical oversight and coaching. The project has exceeded the customer’s expectation and was recommended for adoption in all Gasprom child companies. Two modules developed by our team were honoured with special patents of the Russian Federation.
 \ifFullVersion
 \item Wrote, maintained, and enhanced the numerous smaller projects for Energy, Utility and Oil\&Gas - oriented customers in the Data Capture, Data Verification and ETL areas.
\else
   \item Designed, developed, maintained and enhanced the data collection, verification and migration ETL-like system for Utility and Oil \& Gas - specific databases (both spatial and attribute parts). The changes in the business process proposed by our team and backed by our software enabled a productivity increase in 3-4 times compared to old approach.
\fi
\end{itemize}

\jobdes {PrivateInvest Retail Bank: Kiev, Ukraine}{January 2001 - June 2002}{IT Department Head}
{Retail Bank - specific software}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Assisted in the major expansion of the bank from single office to 22 offices distributed around the country. Created and expanded  the IT department, introduced a number of innovations that allowed us to complete the technical tasks in only 5 months – a month ahead of original schedule.
\end{itemize}

\jobdes {Lost Boys Business Solutions: Amsterdam, The Netherlands}{June 2001 - December 2001}{Software Engineer}
{Java 1.3: JSP, EJB, JDBC, jUnit, Oracle 8i, SQL, PL/SQL, custom-built performance measurement tools}
\begin{itemize} \itemsep -2pt % reduce space between items
 \item Was selected to be part of the special fine-tuning and performance optimization team - a distinction given to the best developers at the Company.
\end{itemize}
\jobdes {KPMG Consulting Barents Group: Kiev, Ukraine}{November 1997 - May 2001}{Analyst Programmer}
{MS SQL Server, MS Office (particularly MS Access), VBA, Delphi, HTML, CSS, JavaScript, Crystal Reports, XML/XSLT}
\begin{itemize}
 \item Planning, design and technical implementation of various projects devoted to the budgeting process of the state and local budgets.  The projects were credited with eliminating 2 months from the budget development cycle and greatly expanded the analytic capabilities of the Ministry of Finance of Ukraine.
\end{itemize}
\jobdes {KiGIS LLC: Kiev, Ukraine / Bratislava, Slovakia}{March 1996 - November 1997}{Software Engineer}
{Geographical Information Systems (GIS)-specific software: Intergraph, Bentley Microstation/Geo graphics}

\osection{EDUCATION}
\jobdes {National Technical University of Ukraine (KPI): Kiev, Ukraine}{Sep 1990 - March 1996}{Engineer Diploma}
 {Engineer Diploma (USSR equivalent of MSc) with Honours in Electrical Engineering}

\jobdes {Coursera and Udacity online courses}{2012 - Present}{Self-study}
{``Functional Programming Principles in Scala'', ``Principles of Reactive Programming (in Scala)'', ``Design and Analysis of Algorithms'' (Parts 1 and 2), ``Introduction to Artificial Intelligence (AI)'', ``Machine Learning (ML)'',``Introduction to Finance'', ``Financial Markets'', ``Model Thinking'', ``Computational Investing'', ``An Introduction to Operations Management'', ``Quantum Mechanics and Quantum Computation'', ``Introduction to Genetics and Evolution''. Currently following ``Data Science'' Specialisation}

\osection{SPECIALS}
\vspace{0.20cm}
\begin{itemize}
 \item Profiles: StackOverflow \url{http://stackoverflow.com/users/900174/l-yara}, GitHub \url{http://github.com/l-yara}, LinkedIn \url{http://uk.linkedin.com/in/ylyakh}.
 \item An active member of Data Science London, London Java Community (LJC), London Scala User Group (LSUG) and London Clojurians communities.
 \item Sports: volleyball (former player of ``CBPhoto'' Irish Division 1 team, now playing in Kent County amateur league), sailing (former pupil of Kiev Olympic School).
\end{itemize}


\end{resume}
\end{document}
