\documentclass{res}

% suppress widows/orphans:
\widowpenalty=1000
\clubpenalty=1000
\usepackage{helvetica} % uses helvetica postscript font
\usepackage{hyperref}
\hypersetup{
    pdfstartview={FitH},   % fits the width of the page to the window
    pdftitle={Resume},
    pdfauthor={Yaroslav Lyakh},
    pdfkeywords={CV} {Java} {multithreading} {RDBMS} {multithreading}, %list of keywords
    colorlinks=true,      % false: boxed links; true: colored links
    % linkcolor=black,          color of internal links (change box color with linkbordercolor)
    % citecolor=green,        color of links to bibliography
    % filecolor=magenta,      color of file links
    urlcolor=black          % color of external links
}
\usepackage{needspace}
\newsectionwidth{0pt}  % So the text is not indented under section headings

\usepackage{scalefnt}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\itemsep=0in
\parsep=0in
\usepackage{enumitem}
\setlist[itemize]{itemsep=-2pt, topsep=4pt, labelindent=\parindent}

% hints to the word-breaking algo
\hyphenation{conti-nuous}

\newcommand{\emp}[1]{{\bf#1}}

\newcommand{\jobdes}[4]{\needspace{3\baselineskip} %to keep with next paragraph
{\noindent \bf #3\hspace{2ex}}{{\em \small #1}}{\hfill #2}\\{{\it \small #4.}}}

\newcommand{\aggjobdes}[4]{\needspace{3\baselineskip} %to keep with next paragraph
{\noindent \bf #1\hspace{2ex}}{\hfill #2}\\
{{\noindent \small \textit{ \textbf{ Companies:} {\hfill #3}}}}\\{{\it \small #4.}}}

\newcommand{\osection}[1]{\section{\sc {\Large \textbf{#1}\\}} \vspace{0.30cm}}

% set formatting, put semicolon and dot as needed
\newcommand{\profileItem}[2]{\item {\sl #1:} #2.}
% \newcommand{\profileItem}[2]{\item {\sc #1:} #2.}

% \sl is slanted, not \it (italic) !!! see also http://www.combinatorics.net/weblib/options/style.html

\usepackage{fancyhdr} % use this package to get a 2 line header
\pagestyle{fancy}     % set pagestyle for document
\renewcommand{\headrulewidth}{0pt} %suppress line drawn by default by fancyhdr
\setlength{\textheight}{10in} %increase text height to fit on 1-page
% \textwidth=5.2in % increase textwidth to get smaller right margin
\setlength{\headsep}{4pt}  % space between header and text
\setlength{\headheight}{12pt} % allow room for header
\fancyheadoffset{1cm}
\rhead{ {\it Yaroslav Lyakh (\href{mailto:l_yara@bigmir.net}{L\_yara@bigmir.net}; tel: +447501 301536)}} %put text in header (right side)
% \rhead{ {\it  \thepage}} put text in header (right side)
\cfoot{}                              %       the foot is empty
\topmargin=-2cm %start text higher on the page
\footskip=4cm

\setlength{\parindent}{2em} % paragraph indent

% switch full / short version
% see http://tex.stackexchange.com/questions/1492/passing-parameters-to-a-document for detail
% use  pdflatex "\def\longversion{1} \input{Yaroslav_Lyakh.tex}"
% command use pandoc for exporting:
% pandoc Yaroslav_Lyakh.tex -o Yaroslav_Lyakh.docx
\begin{document}
% \thispagestyle{empty} %this page has no header
% \name{Yaroslav Lyakh}
\begin{resume}

%\osection{Summary}
% \vspace{0.20cm}
\indent Senior Software Engineer / Senior Java Developer with eighteen years of experience in a variety of roles across a number of industries: Finance, Banking, Web Security, Oil~\&~Gas, Online Gambling, Utilities and Public Safety. More than ten years of Java development experience; extensive involvement in high-load parallel distributed low-latency high-throughput data processing systems development for last five years.

\osection{Technical Profile}
% \vspace{0.20cm}
\begin{itemize}
  \profileItem {Full software development lifecycle (SLDC)} {requirement gathering and analysis, design, prototyping, implementation, performance tuning}
  \profileItem {Development process organisation} {\emp{Scrum} (and other \emp{Agile} flavours), RUP, Waterfall. Team work planning and coordination, peer code reviews, team members coaching and mentoring. \emp{Continuous Integration (CI)} with Jenkins}
  \profileItem {JVM - based languages} {\emp{Java} (from 1.3 to 8.0) SE/EE with \emp{multi-threading, collections} and \emp{synchronisation primitives} (lock/barrier/phaser, Guava Monitor). Basic knowledge of \emp{Scala} and \emp{Groovy}; currently studying \emp{Clojure}}
  \profileItem{Performance diagnosis and optimisation} {performance analysis with \emp{jMeter, VisualVM} and \emp{JMH}. Multi-threading systems development; runtime performance tuning incl. Java GC optimisation}
  \profileItem {Relational Databases (RDBMS)} {\emp{Oracle, MySQL, MS SQL Server;} complex \emp{SQL} query authoring and optimisation; \emp{PL/SQL} stored procedures development and performance tuning}
  \profileItem{Enterprise Technologies} {Enterprise Integration Patterns (EIP), Service Oriented Architecture (SOA), \emp{messaging, caching}. Server-side and middleware development; REST-based microservices implementations; Raft distributed consensus algorithm}
  \profileItem{Data Modelling, Data Analysis and Presentation, Extract-Transform-Load (ETL)} {data denormalisation, complex \emp{data transformation and cleansing}; advanced \emp{report building}}
  \profileItem{Other Languages} {\emp{C\#}, some \emp{Python}, \emp{R/Octave}. Shell Script, XML/XSL, JSON, \LaTeX, MD markup}
  \profileItem{Operation Systems, Tools, Technologies, Libraries} {\emp{Linux, Mac OS} and \emp{Windows}. Apache (Ant, Commons Lang, Commons IO, Log4j, \emp{Maven}, \emp{Tomcat}). Crystal Reports. Google (Guava, \emp{Mockito}). JDBC, Hibernate ORM. jBPM/Activiti. Java Swing. \emp{jUnit}, EasyMock, jMock. \emp{Spring} (Core, Integration, MVC, Roo). IDEs: IntelliJ IDEA, Eclipse, Visual Studio. Source Code Control with \emp{Git} and \emp{SVN/Subversion}. \emp{UML} with Sparx Enterprise Architect and Rational Rose. Emacs}

\end{itemize}

\osection{Career History}
\aggjobdes {Senior Software Engineer / Senior Java Developer}{July 2010 - Present}
{Betfair: 1 year, ION Trading: 1.5 years, Fizzback (NICE Co): 2 years, ScanSafe (Cisco): 1 year (All UK)}
{Low-latency high-throughput systems based on low-level Java 6, 7 and 8. Oracle SQL and PL/SQL. Zookeeper, Raft distributed consensus algorithm. Continuous Integration (CI) with Jenkins and Go CD. TDD (jUnit / TestNG + Mockito), Acceptance Testing (Robot Framework). VisualVM, Groovy, Spring (Core, Spring Integration, SpEL, Spring Roo), Hibernate, Activiti BPM, Maven, Git, Subversion, Scrum, peer reviews, mentoring junior team members}
\begin{itemize}
 \item Developed new functionality for the large-scale, high-throughput modules like One-Way Matching for Betfair, Communication Infrastructure for Fizzback or Data Analysis and Reporting for ScanSafe.
 \item Heavily involved in so-called ``parallelising'' works: redesign and re-implementation of the performance - critical parts as the distributed or at least multithreaded modules: Multiple Orchestrator project in Betfair, Parallel Query Executor in ION Trading and on-the-fly data processing module in Fizzback. While in some cases it was enough to re-code the critical pieces with the “stateless or immutable” approach, most of the existed system required deep redesign or complete re-implementation with the extensive use of Java multi-threading tools.
 \item In Betfair, took a major part in transition of the legacy Oracle-based ``traditional'' (i.e. built on classical read-process-write cycle) Matcher implementation into more robust stream-based architecture based on Raft distributed consensus algorithm. Lead a similar project in Fizzback where we transformed an existing project into much more scalable and reliable pipeline - based solution.
 \item As part of the performance improvement works designed and implemented the project-specific Performance Testing facilities for Betfair, Fizzback and ScanSafe.
 \item Kept a consistent attention on the operational characteristics of the projects. In ION Trading, implemented the Attribute Replacement feature which greatly increased the system’s configuration flexibility and made administering and configuring of the whole product much more convenient. In Fizzback, deeply re-developed the Communication and Reporting sub-modules to make configuration and maintenance more convenient; these changes (like specialised configuration DSL and flexible report templates) greatly improved overall system's maintainability and customer provisioning time.
 \item Enhanced the Continuous Integration infrastructure of all projects I've been involved in. In ION Trading, re-implemented a project-specific Maven plugins and streamlined the existed Robot Framework modules increasing the test reliability and reducing the test execution time. In Fizzback, improved the execution time of the whole Unit Testing process (mostly by database-related optimisation).
 \item Provided training and on-the-fly consultations to the junior team members. In Fizzback, developed good working relations with colleagues from Operations, QA and Customer Support departments; combined with good ``pure'' development work this decreased an average defect response time and improved overall IT team flexibility and confidence.
 \end{itemize}

\aggjobdes {Senior Software Engineer / Principal Software Developer / Architect}{December 2006 - April 2010}
{MapInfo (Pitney Bowes), UK: 1.5 years, Original Solutions, Ireland: 0.5 years, Kring LLC, Ukraine: 1 year}
{.NET (C\#), NUnit, Rhino Mocks, Visual Studio, MS Entity Framework, LINQ, WPF/XAML, Java 5, multi – threading, Web Services, Subversion, Oracle 11g, SQL, PL/SQL, Spring, Hibernate, jBoss jBPM, jUnit, Eclipse, GIS, ETL, TETRA – based mobile devices communication, Scrum}
\begin{itemize} % \itemsep -2pt % reduce space between items
 \item Took a key part in a development of a new MapInfo major product version. Innovative approaches like LINQ and SaaS were featured in this project together with some elements of functional-based design allowing to cut down the codebase size; the stability and reliability of the whole system improved.
 \item Provided the performance optimisation; created a project-specific performance testing subsystem.
 \item Lead the offshore development project of a major Public Safety and Emergency (PS\&E) system development. The project consisted of the server part built on Java (conversion of existing military Command and Control Information System) and mobile .Net Compact Framework - based clients interacting with on-board GPS receivers. A purpose-built integration testing facility was created.
\end{itemize}

\aggjobdes {Team Lead / Software Engineer / Analyst Programmer}{November 1996 - November 2006}
{VNIPITRASNGAZ : 4.5 years, Lost Boys: 0.5 years, KPMG: 3.5 years, KiGIS LLC: 1.5 years (All Ukraine)}
{Java 1.3 and 5, multi-threading, GIS, Oracle Spatial Data Option (SDO), Bentley MicroStation and GeoGraphics, Intergraph suite, Spring, O/R Mapping, Swing, JDBC, jUnit, jMock, Eclipse, Subversion, Oracle 8i, 9i, and 10g, SQL, PL/SQL, UML, Sparx EA, Rational Rose, XML, MS Office + VBA, XSLT, TDD and RUP}
\begin{itemize} % \itemsep -2pt % reduce space between items
 \item Formed and led the Software Development Team involved in the major distributed Geographical Information System (GIS MT) for the Gasprom gas transportation company. The project has exceeded the customer’s expectation and was recommended for adoption in all Gasprom child companies.
   \item Designed, developed, maintained and enhanced the data collection, verification and migration ETL-like system for Utility and Oil \& Gas - specific databases (both spatial and attribute parts). The changes in the business process proposed by our team and backed by our software enabled a productivity increase in 3-4 times compared to old approach.
 \item In Lost Boys, was selected to be part of the special fine-tuning and performance optimization team.
 \item In KPMG, took a first-hand part in a planning, design and implementation of various projects devoted to the budgeting process of the state and local budgets.  The projects were credited with eliminating 2 months from the budget development cycle and greatly expanded the analytic capabilities of the Ministry of Finance of Ukraine.
\end{itemize}

\osection{SPECIALS}
\begin{itemize}
 \item Profiles: StackOverflow \url{http://stackoverflow.com/users/900174/l-yara}, GitHub \url{http://github.com/l-yara}, LinkedIn \url{http://uk.linkedin.com/in/ylyakh}.
 \item Education: Engineer Diploma (USSR equivalent of MSc) in Electrical Engineering from National Technical University of Ukraine (KPI) in Kiev, Ukraine, 1996.
 \item An active member of Data Science London, London Java Community (LJC) and London Clojurians communities. Took a number of online courses on Computer Science (Algorithms, Functional and Reactive Programming); Finance (incl. Financial Markets and Computational Investing); Data Science (incl. Big Data, Machine Learning and Artificial Intelligence) - see full list in LinkedIn profile.
 \item Sports: volleyball (former player of ``CBPhoto'' Irish Division 1 team, now playing in Kent amateur league), sailing (former pupil of Kiev Olympic School).
\item Holding two patents awarded on IT innovations in Data Presentation and Data Visualisation areas.
\item Available from mid-March 2016.
\end{itemize}


\end{resume}
\end{document}
